---
title: "ch6_communities_script.Rmd 04/03/2021"
author: Grant Fleming
output:
  word_document: default
  html_notebook:
    highlight: null
  html_document:
    df_print: paged
---

### Setup 
```{r setup}
# restore packages from lockfile - enter "y" 
# if you are prompted to install or update any packages
renv::restore()

# load critical packages - install if necessary by running
# install.packages(c("here", "pacman"))
library(here)
library(pacman)

# source scripts for our packages and functions
source(here("setup/ch6_packages.R"))
source(here("setup/ch6_functions.R"))

# set a seed
set.seed(10)
```
 
# Compilation

### Load raw Communities and Crime data
```{r}
crime_raw = read_csv(here("data/crimedata.csv"))
```

### View raw data
```{r}
print(crime_raw)
```

### Initial feature selection and reordering
```{r}
crime_raw2 = crime_raw %>% 
  relocate(assaults) %>% 
  select(-c(131:ncol(.))) %>% 
  select(-c(2:6)) %>% 
  mutate(assaults = as.numeric(assaults)) %>% 
  rename_with(tolower) 

```

### Checking for explicit vs. implicit NAs
```{r}
no_na_rows = crime_raw2 %>% 
  drop_na()

na_rows = crime_raw2 %>% 
  anti_join(no_na_rows) 

print(na_rows)
```

### Find proportion of implicit NAs ("?") in each column
```{r}
# how many cols are filled with ? marks
map(crime_raw2, function(x) sum(x == "?")/length(x)) %>% 
  enframe() %>% 
  rename(proportion_missing = value) %>% 
  unnest(proportion_missing) %>% 
  filter(proportion_missing >= 0.10) %>% 
  arrange(-proportion_missing) 
```

### Drop columns with too many implicit NAs ("?")
```{r}
# get a list of the cols that have too many "?"s (above 10%)
drop_cols = map(crime_raw2, function(x) sum(x == "?")/length(x)) %>% 
    enframe() %>% 
    unnest(value) %>% 
    filter(value >= 0.10) %>% 
    pull(name) 

# drop them
crime_raw3 = crime_raw2 %>% 
  select(-all_of(drop_cols))

```

### Check to see if columns add up to 100%
```{r}
crime_raw3 %>% 
  select(contains("racepct")) %>% 
  rowwise() %>% 
  mutate(total = sum(c_across(everything()))) %>% 
  arrange(-total)
```

### Protect groups
```{r}
crime_raw4 = crime_raw3 %>% 
  rowid_to_column() %>% 
  relocate(rowid) %>% 
  select(-matches("percap")) %>% 
  rename_with(., function(x) paste0(x, "_protected"), matches("racepct"))

```

### View protected groups
```{r}
protected_features = crime_raw4 %>% 
  select(matches("_protected")) %>% 
  colnames()

print(protected_features)
```


# Preparation

### Split data
```{r}
crime_split = crime_raw4 %>% 
  initial_split(0.7) # check rules for splitting 

crime_training = training(crime_split)
crime_testing = testing(crime_split)

glimpse(crime_training)
```

### Check correlations among features 
```{r}
crime_features = crime_training %>% 
  select(-c(assaults, rowid))

# function creates one plot per protected feature of the top n = number features that are correlated with
# the particular level of the protected feature (protected group) being plotted. larger gaps between the
# red and black points (average correlation between that feature and all other protected groups) indicate 
# that the feature on the X axis is more correlated with that protected group than with the others. 
check_correlations(crime_training, 
                   drop_cols = c("assaults", "rowid"), 
                   protected_features, 
                   number = 5)
```


### Feature engineering
```{r}
target = "assaults"

form = paste(paste(target, "~"), str_c(colnames(crime_training %>% select(-all_of(target))),
                                         collapse = "+"))

rec = recipe(crime_training, form) %>%
  update_role(matches("_protected"), new_role = "protected") %>%
  step_nzv(all_predictors(), -has_role("protected")) %>%
  step_corr(all_predictors(), threshold = 0.7, -has_role("protected")) %>% 
  step_medianimpute(all_numeric(), -all_outcomes()) %>%
  step_range(all_numeric()) %>%
  prep()

train_data = juice(rec) 
test_data = bake(rec, testing(crime_split))
```

### Show new training data
```{r}
print(train_data)
```

### Create datasheet
```{r}
# forthcoming
```


